Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-02-04 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Move paper to it's own directory Add scripts for figures

We developed a simple open-split interface that delivers a mixture of helium and CO2 to the source at atmospheric pressure.

The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented. A flow of helium is used to displace sample CO2 to the delivery capillary in the split. The core of the design is a dual needle and split housed in a VICI cross. Displacement helium flows in through one side of the cross, into the sample vial through 1/16" stainless steel tubing, which forms the outer flow path of the concentric dual needle. The mixture of sample CO2 and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted(Figure 1). The split has fairly low volume (< 1 mL). Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results.

@@

250cm of 50um capillary defines flow rate from the open split into the source, with a bulkhead connector at the source cage.

@@

Process blank samples were prepared from powdered C-1 or TIRI-F. Standards were prepared using the automated gas handling system by evacuating vials prior to filling with 1 ATM of a modern standard gas.

@@

1.  Displace gas to open split with 100 uL/min He.
2.  Introduce mixture to source \@ 30 uL/min
3.  Acquire data: 90s x 10 measurements.

-   Normalizing standards: live tank gas in vials
-   Blanks: C-1 or TIRI-F in vials

@@

Errors from normalization and blank correction are propagated at each step. The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.

@@

## Data acquisition

A new target was used for each sample. Targets were allowed to "burn in" for 5-10 minutes with pure helium delivered to the source. Initial currents of \> 10 uA dropped rapidly, eventually settling to 100-400 nA. After a stable base current was reached, the sample vial was installed on the double needle, starting sample flow to the source. Sample current rises to a stable level within 30 s, at which time, data acquisition is started. 10 blocks of 90 s for a total of 15 m.

## Data reduction

@@

LBC vs constant contaminant/MBC.

@@

@@

## Reference materials

>>>>

The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented. A flow of helium is used to displace sample CO2 to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross. Displacement helium flows in through one side of the cross, into the sample vial through 1/16" stainless steel tubing, which forms the outer flow path of the concentric dual needle. The mixture of sample CO2 and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1). The split has fairly low volume (< 1 mL). Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results. The system is currently not automated, but may be easily mounted on the arm of a gasbench-style autosampler.

@@

flow rate from the open split into the source is defined by 250cm of 50um glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage. The section inside the cage is jacketed in teflon tubing to help protect against spark damage.

@@

Process blank samples are prepared from powdered C-1 or TIRI-F. Standards are prepared using the automated gas handling system by evacuating vials prior to filling with 1 ATM of a modern standard gas.

@@

Samples are measured on a modified 3MV Tandetron AMS system \@ref(Longworth 2015). The displacement helium is allowed to flow constantly during analysis, and the sample needle is inserted into a helium-filled vial between samples. Measurement of a sample proceeds as follows: a clean sputter target is inserted into the source and the target is sputtered until ten minutes (check this) have passed or blank target current falls below 100nA. The delivery needle is then inserted into the sample vial, which starts sample delivery to the split and source. Currents take ~30s to stabilize, at which point the data acquisition is started. Data is acquired in 10 blocks of 90 s each, for 15 minutes total acquisition time. Normalizing standards, blanks, and secondary standards are measured in the same way. Total time for a measurement, including target pre-cleaning is 20-25 minutes.

Discuss delivery rates and sample sizes here or in results?

Initial currents of \> 10 uA dropped rapidly, eventually settling to 100-400 nA.
1.  Displace gas to open split with 100 uL/min He.
2.  Introduce mixture to source \@ 30 uL/min

@@

Errors from normalization and blank correction are propagated at each step.

@@

@@

#### Current dependent blank

$$I_{sample} + I_{blank} = I_{meas}$$
$$Fm_mI_m=Fm_bI_b+Fm_sI_s$$

$$Fm_s = Fm_m + \frac{I_b(Fm_m-Fm_b)}{I_m-I_b}$$

0.1 μA blank current with 10 μA total = 1% blank
LBC vs constant contaminant/MBC.

@@

#### Current independent blank

We decided to use a simple linear correction. Samples kept within a narrow range of current and to a constant size.

If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.

LBC eqn

$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$

```{r}
# Fix to use current form of processed data
carb_data <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
  filter(!is.na(rec_num),
         Cur > 2) %>% 
  mutate(fm_diff = mean - fm_consensus,
         fm_sigma = sigma(mean, fm_consensus, merr),
         fmbc_diff = fm_lbc - fm_consensus,
         fmbc_sigma = sigma(fm_lbc, fm_consensus, merr))
```

Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`

@@

## Reference materials
The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.
